Adaption of the Chumbley Score to matching of bullet striation marks

### 2018-02-05 doc/paper.Rmd
updating from doc_v2

## Problem statement and motivation

\hh{'considered to be unique' by AFTE - that is right, but we should not use this argument, and I don't think we have to. It's better to focus on the problem at hand: striation marks are used to determine same source for bullets. XXX we could include some figures here XXX Current standards ask for a visual inspection of the marks under a comparison microscope by a firearms examiner. One of the problems raised by the NAS report in 2009 is the subjectivity involved in this comparison and the need for determining error rates \citep{NAS:2009}. There are different approaches used to achieve more objectivity and assess error rates. XXX here we need some examples of approaches XXX
For toolmarks, the Chumbley score has been suggested to determine same-source. XXX summary of results from original paper and the adjustment XXX
In this paper we are interested in how this method has to be adjusted to best be applied to the smaller marks on bullets. }
\gk{done}

Striation marks are used to determine same source for bullets. Current standards ask for a visual inspection of the marks under a comparison microscope by a firearms examiner. One of the problems raised by the NAS report in 2009 is the subjectivity involved in this comparison and the need for determining error rates \citep{NAS:2009}. The issue of subjectivity of a firearms examiner has been reviewed by many authors where things like scientific principle testability and error rates are considered to be defining aspects for an objective analysis such that without identifying the uncertainities associated with a method of comparison, it is inconclusive to regards an analysis complete. This means that with the usual method of visual comparison that is adopted by firearms examiners, lies the problem of lack of any systemized quantification and as such there seems no certain way of associating it with any kind of mathematical probability. Since determining error rates has been duly noted as a fundamental problem in forensic science \citep{NAS:2009}, there is a need of methods that address this problem. Some of the methods that have been used to estimate error rates in firearm examinations include Automatic cartridge case comparison where same source identification of cartridge cases and associated error rates have been provided by \citep{riva}, although as noted by \citep{aoas}, in this case alignments of striae involves roation of planes, which cannot be generalized for bullets. In case of methods that use machine learning algorithms some methods which are bootsrap based methods often tend to give over-estimated error rates \citep{efron}, but there are alternate error estimation techniques as described by \citep{aoas}, \citep{efron} and  \citep{vorburger2016} which give better results.

@@

\citep{chumbley} found that the algorithm gives false-positives in slightly over 2 cases (or 2%) of the time and about 9 false-negatives (or 9%) for every 100 comparisons. \citep{hadler} on the other hand, in the improved version of the algorithm first described by \citep{chumbley} found false-positives for 0 cases and 3 false-negatives (or 6%) for 50 comparisons of knwon matches.
\begin{figure}
\centering
\includegraphics[width=\textwidth]{images/B6-B2-L6.png}
```{r, echo=FALSE, fig.width=8, fig.height=3, out.width='\\textwidth'}
land <- read.csv("../data/b6b2l6.csv")
land %>% ggplot(aes(x = y, y=resid)) +geom_line() +theme_bw()
```

\caption{\label{fig:rgl} Image of a bullet land from a confocal light microscope at 20 fold magnification.}

@@

```{r win-comparison, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap=" The two plots on the left show how the same shift behaves in case of a matching pair and the two plots on the right show how the different shift behaves in case of a matching pair."}
a<- matrix(NA, ncol = 3, nrow = 2)
toolmark.falsepositive<- data.frame(a)
colnames(toolmark.falsepositive)<- c("Classification", "Match", "Non-Match")
toolmark.falsepositive$Classification<- c("Match", "Non-Match")
toolmark.falsepositive$Match<- c(47,0)
toolmark.falsepositive$`Non-Match`<- c(3,50)
toolmark.falsepositive

@@

$V = \dfrac{n_{s}n_{d}}{N\left(N-1\right)}\left[\Sigma^{n_{s}}R_{s}\left(i\right)^{2}+\Sigma^{n_{d}}R_{d}\left(j\right)^{2}\right] -\dfrac{n_{s} n_{d}\left(N+1\right)^{2}}{4\left(N-1\right)}$

## Potential limitations of the Chumbley Score

\hh{in the application to striation marks on bullet lands: smaller in width and curved - need to adjust parameter settings; something similar is done in \citet{afte-chumbley} for toolmark comparisons of slip-joint pliers. XXX work in }

Bullets are much smaller in length, width, are not flat and curved in the cross-sectional topography as opposed to tools like screw driver tips which produced longer and pronounced markings. This means the makings made by barrels in comparison to toolmarks may have a problem in distinctiveness. The majority of Bullet profiles and signatures extracted by procedures mentioned by  \citet{aoas} are almost 1/4 th the size of toolmarks as used by \citet{chumbley} or even smaller. Striations on Bullets are made on their curved surfaces, whereas the algorithm developed by \citet{chumbley} and \citet{hadler} has only been tested for flatter and wider surfaces which have negligible curvature. Therefore, using methods proposed for toolmarks may need adaptation in order to give tangible results for bullets. Moreover, in order to to get flat bullet signatures and remove the curvatures some kind of smoothing needs to be applied as a pre-step which needs further investigation as to whether the level of smoothing does effect the working of the algorithm on Bullets.

Also when in the optimization step, the Window of optimization for bullets will be shorter as the signatures are smaller. The idea is to keep the number of windows of optimization sufficiently large, which means shorter Trace segments (partition of signature or toolmark with length = size of window of optimization) that lets us compare smaller segments of one signature to another.

This introduces a problem as, if we go too small in the window of optimization, the unique features of the trace segments are lost and seem similar, while too large sizes vastly reduces the weight of small features that would otherwise uniquely classify a signature and hence identify the region of agreement.

Thus the Window of Optimization has a direct influence on whether we are
Falsely rejecting the null when it is true (Type I) or Falsely accepting the null when it is false (Type II) making identification of the optimum size of window of optimization very important.

## Testing setup:
\hh{Using cross-validation setup to identify appropriate parameter settings for (a) signatures and (b) profiles directly}

@@

The marking made on bullets are smaller than toolmarks and is also less wider. The idea is to find out possible areas of error while adapting the score based method proposed for toolmarks.

Bullet signatures being compared at this time are from the Hamby 44 and Hamby 252 data, which have the best set of known-matched and known non-matches.

Bullet signatures are chosen by
1. Filtering out Land_id for Profiles from the Hamby 44 and Hamby 252 data and removing 
all NA values
2. run_id = 3 is chosen, Seems like the signatures generated from this run_id give the closest
match. Different run_id's have some different settings for generating the signatures
(The level of smoothing does not seem to be one of them)

The bullet signatures when generated already included the LOWESS smoothing.
Therefore, the coarseness factor is set to 1 while running the chumbley_non_random()
which generates the same_shift, different_shift, U-Stat and P_value parameters.

window Optimization 320 Window Validation 50
```{r echo=FALSE}
all %>% filter(wo==320, wv==50) %>% xtabs(data=., ~signif+match)
```
window Optimization 120 Window Validation 50
```{r echo=FALSE}
all %>% filter(wo==120, wv==50) %>% xtabs(data=., ~signif+match)
```
window Optimization 80 Window Validation 50
```{r echo=FALSE}
all %>% filter(wo==80, wv==50) %>% xtabs(data=., ~signif+match)

@@

# Results

@@

We used the adjusted Chumbley method as proposed in \citet{hadler} and implemented in the R package `toolmaRk` \citep{toolmark} on all pairwise land-to-land comparisons of the Hamby scans (a total of 85,491 comparisons) in the following manner:

```{r, echo=FALSE, results='asis'}
param_setting<- all %>% select(wo, wv) %>% group_by(wo, wv) %>% tally() %>% select(wo,wv)
#param_setting<- data.frame(matrix(param_setting, nrow = 16))
#t(param_setting)
#param_setting %>% tally() 
library(kableExtra)
knitr::kable(t(param_setting), digits=0, caption = "Overview of parameter settings used for optimization and validation windows for bullet land signatures.", format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("scale_down"))
```

Figure \ref{fig:type2} gives an overview of type II error rates observed when varying the window size in the optimization step.

@@

Figure \ref{fig:type1} compares nominal (fixed) type I error and actually observed type I errors for the parameter settings in table XXX. With an increasing size of the window used in the optimization step the observed type I error rate decreases (slighty). This might be related to the increasing number of tests that fail for larger window sizes, in particular for non-matching striae (see fig \ref{fig:missings}).

```{r type1, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap="Comparison of observed and nominal type I error rates  across a range of window sizes for optimization $wo$. The horizontal line in each facet indicates the nominal type I error rate.  As the optimization window increase the observed type I error rate gets smaller. A smaller validation window tends to be associated with a higher type I error rate."}
errors %>% filter(wv %in% c(30,50)) %>%
  ggplot(aes(x = wo, y = actual, colour=factor(alpha))) + 
  geom_hline(aes(yintercept=alpha), colour="grey30") +
  geom_point(aes(shape=factor(wv))) +
  #  facet_grid(.~wo, labeller="label_both") + 
  geom_smooth(aes(group=alpha), se=FALSE, method="lm") +
  theme_bw() +
#  scale_y_log10(breaks=c(0.001,.005, 0.01, 0.05)) +
  scale_shape_discrete("Size of validation window, wv") +
  ylab("Observed type I error rate") +
  xlab(expression("Window size for optimization, "~wo)) +
  scale_colour_brewer(expression("Nominal type I error "~alpha), palette="Set2") +
  facet_wrap(~alpha, labeller="label_both", scales="free") +
  theme(legend.position="bottom")

@@

```{r missings, fig.width=8, fig.height=5, fig.cap="Number of failed tests by the window optimization size, wo, and ground truth. Test results from different sources have a much higher chance to fail, raising the question, whether failed tests should be treated as rejections of the null hypothesis of same source."}
ns <- all %>% group_by(wv, wo, match) %>% summarize(
  n_ss = mean(same_shift_n),
  n_ds = mean(diff_shift_n),
  zero_ss = sum(same_shift_n==0),
  zero_ds = sum(diff_shift_n==0),
  miss = sum(is.na(p_value)),
  missperc = miss/85491*100
)
ns %>% filter(wv %in% c(30, 50)) %>% 
  ggplot(aes(x = wo, y = missperc, shape=factor(wv), colour = factor(match))) + 
  geom_smooth(method="lm", size=.5, se=FALSE) +
  geom_point() + 
  theme_bw() +
  ylab("Percent of failed tests.") +
  xlab("Size of optimization window, wo") +
  scale_colour_brewer("Match", palette="Set1") +
  scale_shape_discrete("Size of \nvalidation\nwindow, wv")
```
Figure \ref{fig:missings} gives an overview of the number of failed tests, i.e. tests in which a particular parameter setting did not return a valid result. This happens, when the shift to align two signatures is so large, that the remaining overlap is too small to accommodate  windows for validation. The problem is therefore exacerbated by a larger validation window. 
Figure \ref{fig:missings} also shows that the number of failed tests is approximately linear in the size of the optimization window.

```{r lms, results='asis'}
nsnest <- ns %>% filter(wv %in% c(30,50)) %>% group_by(match, wv) %>% nest()
nsnest <- nsnest %>% mutate(
  model = data %>% purrr::map(.f = function(d) lm(I(missperc*100)~wo, data=d)),
  coefs = model %>% purrr::map(.f = broom::tidy)
)
tab <- nsnest %>% unnest(coefs) %>% filter(term=="wo") %>% arrange(match) %>%
  select(-term, -statistic, -p.value)
knitr::kable(tab, digits=3, caption = "Estimates of the increase in percent of failed tests corresponding to a 100 point increase in the optimization window.", booktabs=TRUE)
```

```{r wo120, fig.width = 8, fig.height=4.5}
wo120 <- errors %>% filter(wo==120)
wo120 %>% gather(type, error, actual:beta) %>%
  ggplot(aes(x = wv, y = error, colour = factor(alpha)))  + 
  geom_line() + geom_point() + facet_wrap(~type, scales="free") +
  theme_bw()+ 
  theme(legend.position="bottom")
```

Coming back to the database of Hamby-44 and Hamby-252 datasets, the run_id = 3 was used when applying the chumbley algorithm on the signatures. The run_id not only defines the level of smoothing but also signifies the chosen height at which the profiles were selected initially.
Another important aspect is the range of horizontal values (which is referred to as the y values in \citet{aoas}) in the signatures. Thes have already been pre-processed in the database to not include any grooves.

\hh{XXX ideas should go to the previous section on setting up the study; in this section we focus on results only XXX}
The idea therefore is to first use these raw values of the profile directly in the chumbley algorithm, and see how the algorithm performs for different coarseness values (smoothing parameter as referred in the function LOWESS used in the chumbley algorithm)

```{r, fig.cap="Type II errors for different levels of coarseness. XXX We need more values around .25"}
errors2 %>% ggplot(aes(x = coarse, y=beta, colour=factor(alpha))) +
  geom_point() + geom_line() +
  theme_bw() +
  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2")
```

### comparison of total error with signatures and profiles

```{r}
errors %>% ggplot(aes(x = wv, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + facet_grid(~wo) +
  geom_point(data=errors2, aes(shape="profiles"))
```
## Discussion

## Results of Chumbley score
- Nominal alpha value shows dependency on the size of the window of optimization
- Test Fail depends on whether known-match or known non-matches has predictive value
- Type II error least bad for window of validation 30 and window of optimization 120

# Algorithm Modification

The proposed algorithm modification is at the same shift step which comes after the shift distance is identified by finding two windows in the two markings that have the highest correlation.

The modification allows for a "wiggle" room in the second marking of the two sets of markings being compared. This means that each set of same shift windows that are being compared, the second marking will window that is under consideration is allowd to move a little towards the left and a little towards the right.

This gives us a new set of 4 windows, 2 to the left and 2 to the right of the original comparison window. Then the correlations for each one of these 5 windows with the window under consideration of the 1st marking is retrieved (from the validation step correlation matrix) or computed.

Then the window that has the maximum correlation (from the set of 5 windows in the 2nd marking) with the window of the 1st marking is chosen.

This new maximum correlation is then used to compute the U-statistics using the same method as before.

## Expected advantage of modification

We are expecting that this modification would improve the type I and II errors for the good. Less number of false negatives and false positives.
The reason for this expectation is many a times the bullet markings are not made at the same distances for two or more bullets because of the way it comes out of the barrel.

Therefore rigid same shifts might not necessarily compare the right set of windows. In the same shift step for one set of windows, allowing the window in the 2nd marking to wiggle left and right and finding the best match to the window in the 1st marking, lets us adjust for situations where markings are compressed or elongated.

## Dependence on the delta of new windows from the original same shift window

From initial tests the amount of movement to get new windows left and right of the original same shift window seems to directly influence the
the pvalue and U statistic that we look at.

### how much of a movement is reasonable and should be used?
Is there a way to understand if the 2nd marking in comparison to the 1st marking is compressed or elongated or neither? Should the amount of movement selected depend on this compression or elongation? or using a static delta movement value justifieable.

>>>>

## Motivation

Striation marks are used to determine same source for bullets. Current standards ask for a visual inspection of the marks under a comparison microscope by a firearms examiner. One of the problems raised by the NAS report in 2009 is the subjectivity involved in this comparison and the need for determining error rates \citet{NAS:2009}. The issue of subjectivity of a firearms examiner has been reviewed by many authors where things like scientific principle testability and error rates are considered to be defining aspects for an objective analysis such that without identifying the uncertainities associated with a method of comparison, it is inconclusive to regards an analysis complete. This means that with the usual method of visual comparison that is adopted by firearms examiners, lies the problem of lack of any systemized quantification and as such there seems no certain way of associating it with any kind of mathematical probability. Since determining error rates has been duly noted as a fundamental problem in forensic science \citet{NAS:2009}, there is a need of methods that address this problem. Some of the methods that have been used to estimate error rates in firearm examinations include Automatic cartridge case comparison where same source identification of cartridge cases and associated error rates have been provided by \citet{riva}, although as noted by \citet{aoas}, in this case alignments of striae involves roation of planes, which cannot be generalized for bullets. In case of methods that use machine learning algorithms some methods which are bootsrap based methods often tend to give over-estimated error rates \citet{efron}, but there are alternate error estimation techniques as described by \citet{aoas}, \citet{efron} and  \citet{vorburger2016} which give better results.

@@

\citet{chumbley} found that the algorithm gives false-positives in slightly over 2 cases (or 2%) of the time and about 9 false-negatives (or 9%) for every 100 comparisons. \citet{hadler} on the other hand, in the improved version of the algorithm first described by \citet{chumbley} found false-positives for 0 cases and 3 false-negatives (or 6%) for 50 comparisons of knwon matches.
```{r}
a<- matrix(NA, ncol = 3, nrow = 2)
toolmark.falsepositive<- data.frame(a)
colnames(toolmark.falsepositive)<- c("Classification", "Match", "Non-Match")
toolmark.falsepositive$Classification<- c("Match", "Non-Match")
toolmark.falsepositive$Match<- c(47,0)
toolmark.falsepositive$`Non-Match`<- c(3,50)

knitr::kable(toolmark.falsepositive, digits=0, format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("hold"))
```

## Potential limitations of the Chumbley Score adaptation to bullets 
Bullets are much smaller in length, width, are not flat and curved in the cross-sectional topography as opposed to tools like screw driver tips which produced longer and pronounced markings. This means the makings made by barrels in comparison to toolmarks may have a problem in distinctiveness. The majority of Bullet profiles and signatures extracted by procedures mentioned by  \citet{aoas} are almost 1/4 th the size of toolmarks as used by \citet{chumbley} or even smaller. Striations on Bullets are made on their curved surfaces, whereas the algorithm developed by \citet{chumbley} and \citet{hadler} has only been tested for flatter and wider surfaces which have negligible curvature. Therefore, using methods proposed for toolmarks may need adaptation in order to give tangible results for bullets. Moreover, in order to to get flat bullet signatures and remove the curvatures some kind of smoothing needs to be applied as a pre-step which needs further investigation as to whether the level of smoothing does effect the working of the algorithm on Bullets.

Also when in the optimization step, the Window of optimization for bullets will be shorter as the signatures are smaller. The idea is to keep the number of windows of optimization sufficiently large, which means shorter Trace segments (partition of signature or toolmark with length = size of window of optimization) that lets us compare smaller segments of one signature to another. This introduces a problem as, if we go too small in the window of optimization, the unique features of the trace segments are lost and seem similar, while too large sizes vastly reduces the weight of small features that would otherwise uniquely classify a signature and hence identify the region of agreement.

Thus the Window of Optimization has a direct influence on whether we are Falsely rejecting the null when it is true (Type I) or Falsely accepting the null when it is false (Type II) making identification of the optimum size of window of optimization very important. This raises important questions about what parameter settings need to be chosen for bullet land comparison. Something similar is done in \citet{afte-chumbley} for toolmark comparisons of slip-joint pliers where optimum window sizes are determined. There is a need to figure out the best parameter settings which minimizes the errors for unsmoothed markings and pre-processed signatures as determined by \citet{aoas}. An analysis of these error rates and comparison with other methods will help us understand the adaptibility of the chumbley score to bullets.

\hh{figure \ref{fig:rgl} needs to be mentioned somewhere in the write-up}

\begin{figure}
\centering
\includegraphics[width=\textwidth]{images/B6-B2-L6-rescaled.png}
```{r, eval=FALSE, echo=FALSE, fig.width=8, fig.height=3, out.width='\\textwidth'}
land <- read.csv("../data/b6b2l6.csv")
land %>% ggplot(aes(x = y, y=resid)) +geom_line() +theme_bw() +
  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
  xlab(expression(paste("Relative Location (in ",mu,m,")", sep="")))
```

\caption{\label{fig:rgl} Image of a bullet land from a confocal light microscope at 20 fold magnification (top) and a chart of the corresponding signature of the same land (bottom). The dotted lines connect some peaks visible in both visualizations.}

@@

```{r win-comparison, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap=" The two plots on the left show how the same shift behaves in case of a matching pair and the two plots on the right show how the different shift behaves in case of a matching pair."}

@@

$$V = \dfrac{n_{s}n_{d}}{N\left(N-1\right)}\left[\Sigma^{n_{s}}R_{s}\left(i\right)^{2}+\Sigma^{n_{d}}R_{d}\left(j\right)^{2}\right] -\dfrac{n_{s} n_{d}\left(N+1\right)^{2}}{4\left(N-1\right)}$$

## Testing setup

@@

The marking made on bullets are smaller than toolmarks and is also less wider. The idea is to find out possible areas of error while adapting the score based method proposed for toolmarks, using cross-validation setup to identify appropriate parameter settings for (a) signatures and (b) profiles directly

Bullet signatures being compared at this time are therefore from the Hamby 44 and Hamby 252 data. The database setup and pre-processing system used for choosing the Bullet signatures are as described by \citet{aoas}. In order to choose the bullet signatures we first filter out Land_id for Profiles from the Hamby 44 and Hamby 252 data and remove all NA values. Then run_id = 3 is chosen as the signatures generated from this run_id give the closest match. Different run_id's have some different settings for generating the signatures.(The level of smoothing does not seem to be one of them)

The bullet signatures when generated by this process already includes a loess smoothing. Therefore, the coarseness factor is set to 1 while running the chumbley non random algorithm for comparing different optimization windows.The algorithm generates the same_shift, different_shift, U-Stat and P_value parameters which are then used to calculate the errors associated with different sets of window sizes.

Similar to signatures the run_id = 3 was used when applying the chumbley algorithm using the database setup given by \citet{aoas} of Hamby-44 and Hamby-252 datasets, on the profiles. The run_id not only defines the level of smoothing but also signifies the chosen height at which the profiles were selected initially. Another important aspect is the range of horizontal values (which is referred to as the y values in \citet{aoas}) in the signatures. These have already been pre-processed in the database to not include any grooves.

The idea therefore is to first use these raw values of the profile directly in the chumbley algorithm, and see how the algorithm performs for different coarseness values (smoothing parameter as referred in the function lowess used in the chumbley algorithm).

\pagebreak

# Results
We used the adjusted Chumbley method as proposed in \citet{hadler} and implemented in the R package `toolmaRk` \citep{toolmark} on all pairwise land-to-land comparisons of the Hamby scans (a total of 85,491 comparisons) with the pairwise sets for the comparisons given in the table \ref{tab:param}.
```{r param}
param_setting<- all %>% select(wo, wv) %>% group_by(wo, wv) %>% tally() %>% select(wo,wv)
#param_setting<- data.frame(matrix(param_setting, nrow = 16))
#t(param_setting)
#param_setting %>% tally() 
library(kableExtra)
#knitr::kable(t(param_setting)[,1:17], digits=0, caption = "Overview of parameter settings used for optimization and validation windows for bullet land signatures.", format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("scale_down", "hold"))

par_tab1<-t(param_setting)[,1:18]
par_tab2<-t(param_setting)[,19:33]
knitr::kable(list(par_tab1,par_tab2), digits=0,caption = "Overview of parameter settings used for optimization and validation windows for bullet land signatures.", format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("scale_down","hold"))%>% kableExtra::kable_styling(latex_options = c("scale_down","hold")) #%>% group_rows("",1,2) %>% group_rows("",3,4)

@@

```{r confusion}
library(kableExtra)
a27<- all %>% filter(wo==280, wv==30) %>% xtabs(data= ., ~signif+match)
a27<- as.data.frame(a27)
a27$Type<- c("True Negative","False Positive (Type I)","False Negative (Type II)","True Positive")

a28<- all %>% filter(wo==120, wv==30) %>% xtabs(data= ., ~signif+match)
a28<- as.data.frame(a28)
a28$Type<- c("True Negative","False Positive (Type I)","False Negative (Type II)","True Positive")

a29<- all %>% filter(wo==80, wv==30) %>% xtabs(data= ., ~signif+match)
a29<- as.data.frame(a29)
a29$Type<- c("True Negative","False Positive (Type I)","False Negative (Type II)","True Positive")

knitr::kable(list(a27,a28,a29), digits=0, caption = "Confusion Table for different optimization window sizes with validation window size as 30.", format = "latex" ,booktabs=TRUE)%>% kableExtra::kable_styling(latex_options = c("hold"), "striped", full_width = F) %>% group_rows("Size of Optimization Window = 280",1,5) %>% group_rows("Size of Optimization Window = 120", 6, 10) %>%  group_rows("Size of Optimization Window = 80", 11,14)
```

@@

Figure \ref{fig:type2} gives an overview of type II error rates observed when varying the window size in the optimization step. Two levels of validation window size 30 and 50 were chosen as to compare the error rates for different nominal type I errors. We notice that the trends for these nominal type I errors are similar and in most cases a validation window of 50 has higher type II error than for 30. A change in this trend is seen for a 0.05 $\alpha$ level, although the difference between the two windows is very small for this case. We can also notice an obvious trend of increase in the Type II error as the window of optimization increases and see a minimum around the optimization window size of 120 pixels. Hence we are inclined to choose a smaller validation window size and optimization window as 120.

Table \ref{tab:confusion} shows the confusion tables with the classification of type I and type II errors and how the numbers change with a change in the optimization window. The windows represent areas to the left of the window with minimum type II, near the minimum type II window and to the far right of the minimum type II error

@@

Figure \ref{fig:type1} compares nominal (fixed) type I error and actually observed type I errors for the parameter settings in table \ref{tab:param}. With an increasing size of the window used in the optimization step the observed type I error rate decreases (slighty). This means as the optimization window increase the observed type I error rate gets smaller. A smaller validation window on the other hand, tends to be associated with a higher type I error rate. This can be better imagined for a given window of optimization, where the actual Type I error is comparable to the nominal level for only a select few validation window sizes. For these comparable validation window sizes of 30 and 50 as done here, the actual type I error increases very slightly and can be seen in Figure \ref{fig:type1}. This increase is not as much when compared to the variation seen with the optimization window sizes. This effect might be related to the increasing number of tests that fail for larger optimization window sizes, in particular for non-matching striae (see fig \ref{fig:missings}).

```{r type1, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap="Comparison of observed and nominal type I error rates  across a range of window sizes for optimization $wo$. The horizontal line in each facet indicates the nominal type I error rate."}
errors %>% filter(wv %in% c(30,50)) %>%
  ggplot(aes(x = wo, y = actual, colour=factor(alpha))) + 
  geom_hline(aes(yintercept=alpha), colour="grey30") +
  geom_point(aes(shape=factor(wv))) +
  #  facet_grid(.~wo, labeller="label_both") + 
  geom_smooth(aes(group=alpha), se=FALSE, method="lm") +
  theme_bw() +
#  scale_y_log10(breaks=c(0.001,.005, 0.01, 0.05)) +
  scale_shape_discrete("Size of validation window, wv") +
  ylab("Observed type I error rate") +
  xlab(expression("Window size for optimization, "~wo)) +
  scale_colour_brewer(expression("Nominal type I error "~alpha), palette="Set2") +
  facet_wrap(~alpha, labeller="label_both", scales="free") +
  theme(legend.position="bottom")

@@

```{r wo120, fig.width = 4, fig.height=4, fig.cap="The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120"}
wo120 <- errors %>% filter(wo==120)
wo120 %>% gather(type, error, beta) %>%#actual:beta) %>%
  ggplot(aes(x = wv, y = error, colour = factor(alpha)))  +
  geom_line() + geom_point() + facet_wrap(~type, scales="free") +
  theme_bw()+
  theme(legend.position="bottom") +  ggtitle("Signatures" , subtitle = "Varying validation window sizes, Optimization window = 120")
```

The actual type I error and type II error for signatures were also compared for different validation window sizes. Figure \ref{fig:wo120} shows the actual rates for different nominal $\alpha$ levels. We can see that the type II error rises with higher validation windows for the smaller nominal $\alpha$ levels while for the nominal $\alpha$ = 0.05 its almost constant.
```{r }
ns <- all %>% group_by(wv, wo, match) %>% summarize(
  n_ss = mean(same_shift_n),
  n_ds = mean(diff_shift_n),
  zero_ss = sum(same_shift_n==0),
  zero_ds = sum(diff_shift_n==0),
  miss = sum(is.na(p_value)),
  missperc = miss/85491*100
)

nsp <- allp %>% group_by(wv, wo, match, coarse) %>% summarize(
  n_ss = mean(same_shift_n),
  n_ds = mean(diff_shift_n),
  zero_ss = sum(same_shift_n==0),
  zero_ds = sum(diff_shift_n==0),
  miss = sum(is.na(p_value)),
  missperc = miss/85491*100
)

ns_sig_c <- all_sig_c %>% group_by(wv, wo, match, coarse) %>% summarize(
  n_ss = mean(same_shift_n),
  n_ds = mean(diff_shift_n),
  zero_ss = sum(same_shift_n==0),
  zero_ds = sum(diff_shift_n==0),
  miss = sum(is.na(p_value)),
  missperc = miss/85491*100
)

```{r missings, fig.width=6, fig.height=4, fig.cap="Number of failed tests by the window optimization size, wo, and ground truth.",fig.align = "center"}
#ns1<- ns[which(ns$wv == c(30,50), arr.ind = T),]
missings<- ns %>% filter(wv %in% c(30, 50)) %>%
#ns %>% filter(wv == c(30, 50)) %>%
  ggplot(aes(x = wo, y = missperc, shape=factor(wv), colour = factor(match)), size=.3) + 
  geom_smooth(method="lm", size=.2, se=FALSE) +
  geom_point() + 
  theme_bw() +
  ylab("Percent of failed tests.") +
  xlab("Size of optimization window, wo") +
  scale_colour_brewer("Match", palette="Set1") +
  scale_shape_discrete("Size of \nvalidation\nwindow, wv")
missings
```

```{r lms, results='asis'}
nsnest <- ns %>% filter(wv %in% c(30,50)) %>% group_by(match, wv) %>% nest()
nsnest <- nsnest %>% mutate(
  model = data %>% purrr::map(.f = function(d) lm(I(missperc*100)~wo, data=d)),
  coefs = model %>% purrr::map(.f = broom::tidy)
)
tab <- nsnest %>% unnest(coefs) %>% filter(term=="wo") %>% arrange(match) %>%
  select(-term, -statistic, -p.value)
knitr::kable(tab, digits=3, caption = "Estimates of the increase in percent of failed tests corresponding to a 100 point increase in the optimization window.", booktabs=TRUE)
```
Figure \ref{fig:missings} gives an overview of the number of failed tests, i.e. tests in which a particular parameter setting did not return a valid result. This happens, when the shift to align two signatures is so large, that the remaining overlap is too small to accommodate  windows for validation. The problem is therefore exacerbated by a larger validation window. Figure \ref{fig:missings} also shows that the number of failed tests is approximately linear in the size of the optimization window. Test results from different sources have a much higher chance to fail, raising the question, whether failed tests should be treated as rejections of the null hypothesis of same source. For known non-matches there is a higher possibility that in the optimization pair of windows where cross-correlations are maximum are too far apart, and same shifts of this order hit the end of the signature.

## Profiles

Figure \ref{fig:prof_missings} (a) shows the type II error rates for profiles for the optimization window 120 and validation window 30 with varying level of coarseness. We can see that the type II error for all the nominal $\alpha$ levels is lowest in the range of 0.20 to 0.35. Therefore, a value of 0.25 can be used keeping in mind it keeps the type II error lowest while runnning simulations. Thus for comparisons of different window sizes etc as seen in the differnt parts of Figure \ref{fig:prof_missings} this coarseness value is used.

On the other hand Figure \ref{fig:prof_missings} (b) shows if the coarseness level set in the chumbley agorithm has any effect on the signatures, which are pre-processed and already smoothed to a certain extent. From Figure \ref{fig:prof_missings} (b) we can notice that for different nominal $\alpha$ levels, the type II error fluctuates slightly but does not change much, thereby helping us conclude that the coarseness levels set in the LOWESS smoothing in the chumbley alggorithm does effect the type II error much for signatures.

### Comparison of profiles and signatures

Another reason for failed tests can be incorrect identification of maximum correlation windows in the optimization step as seen in figure \ref{fig:prof_missings}(d) because of the level of smoothing, as too much smoothing would subdue intricate features that might otherwise help in the correlation calculations and correct identification of maximum correlation windows irrespective of the size. This would again cause a simiar effect as explained for figure \ref{fig:missings} with validation windows, irrespective of size, during the shifts end up at the ends of the markings resulting in an invalid calculation and failed comparison attempt.

In figure \ref{fig:prof_missings}(d) and (f), we compare profiles and signatures on the basis of number of failed tests. The profiles chosen for figure \ref{fig:prof_missings}(f) have a constant coarseness of 0.25 and window of optimization as 120. The signatures in this case are not smoothed using the chumbley algorithm step of LOWESS smoothing. Instead signatures are used as calculated by \citet{aoas}. The smoothing in these signatures were determined and fixed on the basis of their performance in the random forest based algorithm proposed by \citet{aoas}. The comparison of profiles and signatures with variation of validation window size therefore is made on even footing. The trends are similar to figure \ref{fig:missings} in the sense that for known non-matches the number of failed tests are more for both signatures and profiles and increasing linearly with the validation window size. The problem is however, worse for profiles which has higher number of failed tests than signatures for all validation windows.

The total error for different validation window sizes for signatures and profiles can be seen in figure \ref{fig:prof_missings} (e).The optimization window size is 120 and profiles are calculated at a default 0.25 coarseness level while signatures as before are not smoothed again in the modified chumbley algorithm. We can see that the total error is always higher for profiles as compared to signatures for all sizes of validation window.

```{r, fig.cap="Type II errors for different levels of coarseness."}
e1<-errors2 %>% filter(wv==30)%>% ggplot(aes(x = coarse, y=beta, colour=factor(alpha))) +
  geom_point() + geom_line() +
  theme_bw() +
  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + xlab("Profile Coarseness")+
  ylab("Type II error") + ggtitle(label = "(a)")

e2<- errors_sig_c %>% filter(wv==30)%>% ggplot(aes(x = coarse, y=beta, colour=factor(alpha))) +
  geom_point() + geom_line() +
  theme_bw() +
  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2")+ xlab("Signature Coarseness") + ylab("Type II error") + ggtitle(label = "(b)")

#multiplot(e1,e2, cols = 1)

```{r}
#errors %>% ggplot(aes(x = wv, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + facet_grid(~wo) +geom_point(data=errors2, aes(shape="profiles"))

total1<- errors_sig_c %>% ggplot(aes(x = coarse, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + geom_line()  + geom_point(data= errors2 %>% filter(wv==30), aes(shape="profiles")) + geom_line(data=errors2 %>% filter(wv==30), aes(shape="profiles")) + theme_bw() +
  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + xlab("Coarseness") +ylab("Total error") + ggtitle(label = "(c)")

total2<- wo120 %>% filter(wv>10) %>% ggplot(aes(x = wv, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + geom_line() +geom_point(data=errors2 %>% filter(wv>10) %>% filter(coarse==0.25), aes(shape="profiles")) + geom_line(data=errors2 %>% filter(wv>10)%>% filter(coarse==0.25), aes(shape="profiles")) + theme_bw() +
  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + ylab("Total error") + ggtitle(label = "(e)")
